---
title: "Modeling 5"
editor: visual
author: Emi Cervantes
toc: true
date: "9-13-2023"
number-sections: true
highlight-style: pygments
format: 
  html: 
    code-fold: false
    html-math-method: katex
    fig-width: 7
    fig-height: 4
  pdf:
    toc: true
    number-sections: true
    geometry:
      - top=30mm
      - left=20mm
    fig-width: 7
    fig-height: 4
  docx: 
    fig-width: 10
    fig-height: 5
---

**Modeling Procedures**

1)  Fulfill the requirements to perform a path analysis: assumptions regarding sample size, multivariate normal distribution, and multicollinearity

## Load Libraries

```{r}
library(tidyverse)
library(lavaan)
library(semPlot)
```

## Load Data

```{r}
df <- read_csv("../../../data/df_imputed.csv")
head(df)
```

### Data Clearning

We need to add average for situational interest and mind wandering

```{r}
# average for situational interest and mind wandering
df <- df %>% mutate(SI_avg = rowMeans(df[12:25], na.rm=TRUE),
              MW_avg = rowMeans(df[26:32], na.rm=TRUE))
head(df[41:42])
```

```{r}
# Standardize data
df2 <- df %>% select(Sex, Condition, chicago,
                           TMA_avg, SI_avg, MW_avg, Del_OverallAcc, Understand_avg)
df2[4:8] <- df2[4:8] %>% mutate_all(~(scale(.) %>% as.vector))
```

```{r}
# subgroup data
df_we <-  df2 %>% filter(Condition == 1)
# chicago
df_we_c <-  df_we %>% filter(chicago == 1)
df_we_i <-  df_we %>% filter(chicago == 0)
# irvine
df_nwe <-  df2 %>% filter(Condition == 0)
# chicago
df_nwe_c <-  df_nwe %>% filter(chicago == 1)
# irvine
df_nwe_i <-  df_nwe %>% filter(chicago == 0)
```

## Model

### Hypotheses

$H1$: Trait math anxiety will have a direct positive effect on mind wandering and will have a direct negative effect on learning achievement

$H2$: Situational interest will have a direct positive effect on mind wandering, and will have a direct positive effect on learning achievement

$H3$: Mind wandering has a direct negative effect on learning achievement

$H4$: Trait math anxiety will have an indirect negative effect on learning achievement

$H5$: Situational interest will have an indirect positive effect on learning achievement

```{r}
# covariance table
df2[4:8] %>% cov(use = "complete.obs")
```

### Overall Accuracy Scores

#### Model Definition

```{r}
# model definition
m_acc <- '
# regressions
MW_avg ~ 1 + TMA_avg + SI_avg
Del_OverallAcc ~ 1 + TMA_avg + MW_avg + SI_avg
'
```

#### Worked EX Group

##### Full Sample

```{r}
fit1_we <- sem(m_acc, data = df_we, 
            std.lv = T, group = "Sex")
fit2_we <- sem(m_acc, data = df_we, group = "Sex",
            group.equal = "intercepts")
fit3_we <- sem(m_acc, data = df_we, group = "Sex",
            group.equal = c("intercepts", "regressions"))
summary(fit1_we)
lavTestLRT(fit1_we, fit2_we, fit3_we)
```

Visualization:

```{r}
semPaths(fit1_we,
         whatLabels = "est",
         sizeMan = 10,
         edge.label.cex = 1.15,
         style = "ram",
         nCharNodes = 0, nCharEdges = 0,
         nodeLabels = c("MW", "Accuracy",
                        "TMA","SI"),
         intercepts = FALSE)
```

**Results**

Trait math anxiety had a positive direct effect on mind wandering for girls ($\beta = 0.597, p < 0.0001$). Boys did not have any direct effect on mind wandering by any predictors.

None of the predictors had any significant effects on students' overall accuracy scores for boys and girls.

##### Chicago

```{r}
fit1_we_c <- sem(m_acc, data = df_we_c, 
            std.lv = T, group = "Sex")
fit2_we_c <- sem(m_acc, data = df_we_c, group = "Sex",
            group.equal = "intercepts")
fit3_we_c <- sem(m_acc, data = df_we_c, group = "Sex",
            group.equal = c("intercepts", "regressions"))
summary(fit1_we_c)
lavTestLRT(fit1_we_c, fit2_we_c, fit3_we_c)
```

Visualization:

```{r}
semPaths(fit1_we_c,
         whatLabels = "est",
         sizeMan = 10,
         edge.label.cex = 1.15,
         style = "ram",
         nCharNodes = 0, nCharEdges = 0,
         nodeLabels = c("MW", "Accuracy",
                        "TMA","SI"),
         intercepts = FALSE)
```

**Results**

Trait math anxiety had a positive direct effect on mind wandering ($\beta = 0.847, p < 0.001$) for girls. On the other hand, for boys, neither trait math anxiety or situational interests had any direct effects on mind wandering.

None of the predictors had any significant effects on girls' overall accuracy scores. For boys, trait math anxiety had negative direct effect on accuracy scores ($\beta = -2.090, p = 0.001$) and situational interest also had negative direct effects on accuracy scores. ($\beta = -2.226, p = 0.002$)

##### Irvine

```{r}
fit1_we_i <- sem(m_acc, data = df_we_i, 
            std.lv = T, group = "Sex")
fit2_we_i <- sem(m_acc, data = df_we_i, group = "Sex",
            group.equal = "intercepts")
fit3_we_i <- sem(m_acc, data = df_we_i, group = "Sex",
            group.equal = c("intercepts", "regressions"))
summary(fit1_we_i)
lavTestLRT(fit1_we_i, fit2_we_i, fit3_we_i)
```

Visualization:

```{r}
semPaths(fit1_we_c,
         whatLabels = "est",
         sizeMan = 10,
         edge.label.cex = 1.15,
         style = "ram",
         nCharNodes = 0, nCharEdges = 0,
         nodeLabels = c("MW", "Accuracy",
                        "TMA","SI"),
         intercepts = FALSE)
```

**Results**

Girls had no direct effects on mind wandering. For boys, trait math anxiety had positive direct effect ($\beta = 0.338, p = 0.021$) on mind wandering.

Girls did not have any direct/indirect effects on accuracy scores. Boys had both positive effects by trait math anxiety ($\beta = 0.260, p = 0.052$) and situational interest ($\beta = 0.477, p = 0.0001$)

#### No Wokred EX Group

##### Full Sample

```{r}
fit1_nwe <- sem(m_acc, data = df_nwe, group = "Sex")
fit2_nwe <- sem(m_acc, data = df_nwe, group = "Sex",
                group.equal = "intercepts")
fit3_nwe <- sem(m_acc, data = df_nwe, group = "Sex",
                group.equal = c("intercepts", "regressions"))
summary(fit1_nwe)
lavTestLRT(fit1_nwe, fit2_nwe, fit3_nwe)
```

Visualization:

```{r}
semPaths(fit1_nwe,
         whatLabels = "est",
         sizeMan = 10,
         edge.label.cex = 1.15,
         style = "ram",
         nCharNodes = 0, nCharEdges = 0,
         nodeLabels = c("MW", "Accuracy",
                        "TMA","SI"),
         intercepts = FALSE)
```

**Results**

Girls had positive direct effect on mind wandering by trait math anxiety ($\beta = 0.375, p = 0.0001$). Boys did not have any direct effect on mind wandering.

Girls had negative direct effect on accuracy scores by trait math anxiety ($\beta = -0.332, p = 0.029$). Boys had negative direct on accuracy scores by trait math anxiety ($\beta = -0.273, p = 0.044$).

Girls had worst effect on trait math anxiety than boys.

##### Chicago

```{r}
fit1_nwe_c <- sem(m_acc, data = df_nwe_c, 
            std.lv = T, group = "Sex")
fit2_nwe_c <- sem(m_acc, data = df_nwe_c, group = "Sex",
            group.equal = "intercepts")
fit3_nwe_c <- sem(m_acc, data = df_nwe_c, group = "Sex",
            group.equal = c("intercepts", "regressions"))
summary(fit1_nwe_c)
lavTestLRT(fit1_nwe_c, fit2_nwe_c, fit3_nwe_c)
```

Visualization:

```{r}
semPaths(fit1_nwe_c,
         whatLabels = "est",
         sizeMan = 10,
         edge.label.cex = 1.15,
         style = "ram",
         nCharNodes = 0, nCharEdges = 0,
         nodeLabels = c("MW", "Accuracy",
                        "TMA","SI"),
         intercepts = FALSE)
```

**Results**

Girls did not have any direct effects on mind wandering. Boys had direct positive effects on mind wandering by both trait math anxiety ($\beta = 0.419, p = 0.011$) and situational interest ($\beta = 0.487, p = 0.020$).

Girls did not have any direct effects on accuracy scores. Boys had significant negative direct effect on accuracy scores by trait math anxiety ($\beta = -0.580, p = 0.040$).

##### Irvine

```{r}
fit1_nwe_i <- sem(m_acc, data = df_nwe_i, 
            std.lv = T, group = "Sex")
fit2_nwe_i <- sem(m_acc, data = df_nwe_i, group = "Sex",
            group.equal = "intercepts")
fit3_nwe_i <- sem(m_acc, data = df_nwe_i, group = "Sex",
            group.equal = c("intercepts", "regressions"))
summary(fit1_nwe_i)
lavTestLRT(fit1_nwe_i, fit2_nwe_i, fit3_nwe_i)
```

Visualizations:

```{r}
semPaths(fit1_nwe_i,
         whatLabels = "est",
         sizeMan = 10,
         edge.label.cex = 1.15,
         style = "ram",
         nCharNodes = 0, nCharEdges = 0,
         nodeLabels = c("MW", "Accuracy",
                        "TMA","SI"),
         intercepts = FALSE)
```

**Results**

Girls did not have any direct effects on mind wandering. Boys had positive direct effect on mind wandering by trait math anxiety ($\beta = 0.390, p = 0.0001$).

Girls did not have any direct effects on accuracy scores. Boys had a negative direct effect on overall accuracy by trait math anxiety ($\beta = -0.381, p = 0.003$).

### Perceived Understanding

#### Model Definition

```{r}
# model definition
m_und <- '
# regressions
MW_avg ~ 1 + TMA_avg + SI_avg
Understand_avg ~ 1 + TMA_avg + MW_avg + SI_avg
'
```

#### Worked EX Group

##### Full Sample

```{r}
fit1_we <- sem(m_und, data = df_we, 
            std.lv = T, group = "Sex")
fit2_we <- sem(m_und, data = df_we, group = "Sex",
            group.equal = "intercepts")
fit3_we <- sem(m_und, data = df_we, group = "Sex",
            group.equal = c("intercepts", "regressions"))
summary(fit1_we)
lavTestLRT(fit1_we, fit2_we, fit3_we)
```

Visualization:

```{r}
semPaths(fit1_we,
         whatLabels = "est",
         sizeMan = 10,
         edge.label.cex = 1.15,
         style = "ram",
         nCharNodes = 0, nCharEdges = 0,
         nodeLabels = c("MW", "Understanding",
                        "TMA","SI"),
         intercepts = FALSE)
```

**Results**

Girls had a positive direct effect on mind wandering by trait math anxiety ($\beta = 0.597, p = 0.0001$). Boys did not have any significant direct effect on mind wandering.

Girls did not have any direct effects on perceived understanding. Boys had a positive direct effect on perceived understanding by situational interest ($\beta = 0.351, p = 0.001$).

##### Chicago

```{r}
fit1_we_c <- sem(m_und, data = df_we_c, 
            std.lv = T, group = "Sex")
fit2_we_c <- sem(m_und, data = df_we_c, group = "Sex",
            group.equal = "intercepts")
fit3_we_c <- sem(m_und, data = df_we_c, group = "Sex",
            group.equal = c("intercepts", "regressions"))
summary(fit1_we_c)
lavTestLRT(fit1_we_c, fit2_we_c, fit3_we_c)
```

Visualizations:

```{r}
semPaths(fit1_we_c,
         whatLabels = "est",
         sizeMan = 10,
         edge.label.cex = 1.15,
         style = "ram",
         nCharNodes = 0, nCharEdges = 0,
         nodeLabels = c("MW", "Understanding",
                        "TMA","SI"),
         intercepts = FALSE)
```

**Results**

Girls had a positive direct effect on mind wandering by trait math anxiety ($\beta = 0.854, p = 0.001$). Boys did not have any direct effect on mind wandering.

Girls did not have any direct effects on perceived understanding. Boys had a positive effect on perceived understanding by situational interest ($\beta = 1.246, p = 0.008$).

##### Irvine

```{r}
fit1_we_i <- sem(m_und, data = df_we_i, 
            std.lv = T, group = "Sex")
fit2_we_i <- sem(m_und, data = df_we_i, group = "Sex",
            group.equal = "intercepts")
fit3_we_i <- sem(m_und, data = df_we_i, group = "Sex",
            group.equal = c("intercepts", "regressions"))
summary(fit1_we_i)
lavTestLRT(fit1_we_i, fit2_we_i, fit3_we_i)
```

Visualizations:

```{r}
semPaths(fit1_we_i,
         whatLabels = "est",
         sizeMan = 10,
         edge.label.cex = 1.15,
         style = "ram",
         nCharNodes = 0, nCharEdges = 0,
         nodeLabels = c("MW", "Understanding",
                        "TMA","SI"),
         intercepts = FALSE)
```

**Results**

Girls did not have any direct effect on mind wandering. Boys had a positive direct effect on mind wandering by trait math anxiety ($\beta = 0.338, p = 0.021$).

Girls had a positive effect on perceived understanding by situational interest ($\beta = 0.368, p = 0.001$). Boys had a positive effect on perceived understanding by situational interest ($\beta = 0.387, p = 0.003$).

Effects on perceived understanding by situational interest were somewhat the same for both girls and boys.

#### No Wokred EX Group

##### Full Sample

```{r}
fit1_nwe <- sem(m_und, data = df_nwe, group = "Sex")
fit2_nwe <- sem(m_und, data = df_nwe, group = "Sex",
                group.equal = "intercepts")
fit3_nwe <- sem(m_und, data = df_nwe, group = "Sex",
                group.equal = c("intercepts", "regressions"))
summary(fit1_nwe)
lavTestLRT(fit1_nwe, fit2_nwe, fit3_nwe)
```

Visualizations:

```{r}
semPaths(fit1_nwe,
         whatLabels = "est",
         sizeMan = 10,
         edge.label.cex = 1.15,
         style = "ram",
         nCharNodes = 0, nCharEdges = 0,
         nodeLabels = c("MW", "Understanding",
                        "TMA","SI"),
         intercepts = FALSE)
```

**Results**

Girls had a positive direct effect on mind wandering by trait math anxiety ($\beta = 0.375, p = 0.0001$). Boys had no direct effect on mind wandering.

Girls did not have any direct effects on perceived understanding. Boys had a negative direct effect on perceived understanding by mind wandering ($\beta = -0.342, p = 0.010$).

##### Chicago

```{r}
fit1_nwe_c <- sem(m_und, data = df_nwe_c, 
            std.lv = T, group = "Sex")
fit2_nwe_c <- sem(m_und, data = df_nwe_c, group = "Sex",
            group.equal = "intercepts")
fit3_nwe_c <- sem(m_und, data = df_nwe_c, group = "Sex",
            group.equal = c("intercepts", "regressions"))
summary(fit1_nwe_c)
lavTestLRT(fit1_nwe_c, fit2_nwe_c, fit3_nwe_c)
```

Visualizations:

```{r}
semPaths(fit1_nwe_c,
         whatLabels = "est",
         sizeMan = 10,
         edge.label.cex = 1.15,
         style = "ram",
         nCharNodes = 0, nCharEdges = 0,
         nodeLabels = c("MW", "Understanding",
                        "TMA","SI"),
         intercepts = FALSE)
```

**Results**

Girls did not have any direct effect on mind wandering. Boys had positive effects on perceived understanding by both trait math anxiety ($\beta = 0.419, p = 0.011$) and situational interest ($\beta = 0.487, p = 0.020$).

Girls did not have any direct effect on perceived understanding. Boys had a negative direct effect on perceived understanding by trait math anxiety ($\beta = -0.580, p = 0.040$).

##### Irvine

```{r}
fit1_nwe_i <- sem(m_und, data = df_nwe_i, 
            std.lv = T, group = "Sex")
fit2_nwe_i <- sem(m_und, data = df_nwe_i, group = "Sex",
            group.equal = "intercepts")
fit3_nwe_i <- sem(m_und, data = df_nwe_i, group = "Sex",
            group.equal = c("intercepts", "regressions"))
summary(fit1_nwe_i)
lavTestLRT(fit1_nwe_i, fit2_nwe_i, fit3_nwe_i)
```

Visualizations:

```{r}
semPaths(fit1_nwe_i,
         whatLabels = "est",
         sizeMan = 10,
         edge.label.cex = 1.15,
         style = "ram",
         nCharNodes = 0, nCharEdges = 0,
         nodeLabels = c("MW", "Understanding",
                        "TMA","SI"),
         intercepts = FALSE)
```

**Results**

Girls had no direct effect on mind wandering. Boys had a positive direct effect on mind wandering by trait math anxiety ($\beta = 0.390, p = 0.000$).

Girls had no direct/indirect effect on perceived understanding. Boys had a negative direct effect on perceived understanding by trait math anxiety ($\beta = -0.381, p = 0.003$).
