---
title: "Path Analysis w/ Moderation Effect"
format: html
editor: visual
---

# Path Analysis w/ Moderation Effect

## Load Libraries

```{r}
library(tidyverse)
library(lavaan)
library(semPlot)
```

## Load Data

```{r}
df <- readxl::read_xlsx('../../../data/math-anxiety-raw-data.xlsx')
head(df)
```

### Data Clearning

```{r}
tma_lst <- c("TMA_1", "TMA_2", "TMA_3","TMA_4", "TMA_5", "TMA_6",
             "TMA_sum", "TMA_avg")
df2 <- df %>% select(Condition, Sex, chicago, nonwhite, pretest,
              MW_day1_avg, MW_day2_avg, SI_avg, tma_lst,
              Understand_avg, Del_OverallAcc)
# Get rid of row that has NA values in Condition and/or Sex
df2 <- df2 %>% filter(Condition == 1 | Condition == 2)
df2 <- df2 %>% filter(!is.na(Sex))
df2$Condition[df2$Condition == 2] <- 0
# Assign 0 for "boy" and 1 for "girl"
df2$Sex[df2$Sex == 1] <- 0
df2$Sex[df2$Sex == 2] <- 1
# Make sure variable is categorical
df2$chicago <- as.numeric(df2$chicago)
df2$Sex <- as.numeric(df2$Sex)
df2$nonwhite <- as.numeric(df2$nonwhite)
df2$Condition <- as.numeric(df2$Condition)
```

Add MW average

```{r}
df2 <- df2 %>% mutate(MW_avg = rowMeans(df2[6:7], na.rm=TRUE))
head(df2[19])
```

Standardize data

```{r}
df2 <- df2 %>% select(Sex, Condition, chicago,
                           TMA_avg, SI_avg, MW_avg, Del_OverallAcc, Understand_avg)
df2[4:8] <- df2[4:8] %>% mutate_all(~(scale(.) %>% as.vector))
```

## Modeling

### Overall Accuracy

**Model definition**

```{r}
m_acc <- '
# regressions
MW_avg ~ 1 + TMA_avg + SI_avg + Sex + TMA_avg:Sex
Del_OverallAcc ~ 1 + TMA_avg + MW_avg + SI_avg + Sex + MW_avg:Sex + TMA_avg:Sex
'
```

**Implement model**

Fitting the model to entire dataset:

```{r}
fit_acc1 <- sem(m_acc, data = df2, std.lv = T)
summary(fit_acc1)
```

Multigroup analysis with condition:

```{r}
fit_acc2 <- sem(m_acc, data = df2, std.lv = T, group = "Condition")
summary(fit_acc2)
```

Looking closely at the differences in groups

```{r}
fit_acc2.constrained <- sem(m_acc, df2, group = "Condition", 
    group.equal = c("intercepts", "regressions"))
anova(fit_acc2, fit_acc2.constrained)
```

Some paths in the model are different among groups -\> Looking closely at the differences in the effect of math anxiety on mind wandering between groups

```{r}
# redefine model
m_acc.constrained <- '
# regressions
MW_avg ~ 1 + c("b1", "b1") * TMA_avg + SI_avg + Sex +  TMA_avg:Sex
Del_OverallAcc ~ 1 + TMA_avg + MW_avg + SI_avg + Sex + MW_avg:Sex +  TMA_avg:Sex
'
# fit model
fit_acc2.constrained2 <- sem(m_acc.constrained, df2, group = "Condition")
anova(fit_acc2, fit_acc2.constrained2)
```

No significant difference between worked ex groups vs no worked ex groups on the effect of math anxiety on mind wandering.

### Perceived Understanding

**Model definition**

```{r}
m_und <- '
# regressions
MW_avg ~ 1 + TMA_avg + SI_avg + Sex + TMA_avg:Sex
Understand_avg ~ 1 + TMA_avg + MW_avg + SI_avg + Sex + MW_avg:Sex + TMA_avg:Sex
'
```

**Implement model**

Fitting the model to entire dataset:

```{r}
fit_und1 <- sem(m_und, data = df2, std.lv = T)
summary(fit_und1)
```

Multigroup analysis with condition:

```{r}
fit_und2 <- sem(m_und, data = df2, std.lv = T, group = "Condition")
summary(fit_und2)
```

Finding model difference between groups:\

```{r}
fit_und2.constrained <- sem(m_und, df2, group = "Condition", 
    group.equal = c("intercepts", "regressions"))
anova(fit_und2, fit_und2.constrained)
```

Some paths in the model may vary among groups -\> Looking at the effect of mind wandering on perceived understanding

```{r}
# redefine model
m_und.constrained <- '
# regressions
MW_avg ~ 1 + TMA_avg + SI_avg + Sex +  TMA_avg:Sex
Understand_avg ~ 1 + TMA_avg + c("b1", "b1") * MW_avg + SI_avg + Sex + MW_avg:Sex +  TMA_avg:Sex
'
# fit model
fit_und2.constrained2 <- sem(m_und.constrained, df2, group = "Condition")
anova(fit_und2, fit_und2.constrained2)
```

No significant difference was observed on the effect of mind wandering on perceived understanding.

## Modeling w/ Imputed Data

### Load Dataset

```{r}
df_imput <- read_csv("../../../data/df_imputed.csv")
head(df_imput)
```

Clean Data

```{r}
# average for situational interest and mind wandering
df_imput <- df_imput %>% 
  mutate(SI_avg = rowMeans(df_imput[12:25], na.rm=TRUE),
              MW_avg = rowMeans(df_imput[26:32], na.rm=TRUE))
head(df_imput[41:42])
```

```{r}
# Standardize data
df_imput <- df_imput %>% select(Sex, Condition, chicago,
                           TMA_avg, SI_avg, MW_avg, Del_OverallAcc, Understand_avg)
df_imput[4:8] <- df_imput[4:8] %>% mutate_all(~(scale(.) %>% as.vector))
```

### Modeling

#### Overall Accuracy

**Implement model**

Fitting the model to entire dataset:

```{r}
fit_acc1_2 <- sem(m_acc, data = df_imput, std.lv = T)
summary(fit_acc1_2)
```

Multigroup analysis with condition:

```{r}
fit_acc2_2 <- sem(m_acc, data = df_imput, std.lv = T, 
                  group = "Condition")
summary(fit_acc2_2)
```

#### Perceived Understanding

**Implement model**

Fitting the model to entire dataset:

```{r}
fit_und2_1 <- sem(m_und, data = df_imput, std.lv = T)
summary(fit_und2_1)
```

Multigroup analysis with condition:

```{r}
fit_und2_2 <- sem(m_und, data = df_imput, std.lv = T, 
                  group = "Condition")
summary(fit_und2_2)
```
