---
title: "Path Analysis w/ Moderation Effect"
editor: visual
author: Emi Cervantes
toc: true
date: ""
number-sections: true
highlight-style: pygments
format: 
  html: 
    code-fold: false
    html-math-method: katex
    fig-width: 7
    fig-height: 4
  pdf:
    toc: true
    number-sections: true
    geometry:
      - top=30mm
      - left=20mm
    fig-width: 7
    fig-height: 4
  docx: 
    fig-width: 10
    fig-height: 5
---

# Path Analysis w/ Moderation Effect

## Load Libraries

```{r}
library(tidyverse)
library(lavaan)
library(semPlot)
library(skimr)
library(janitor)
```

## Load Data

```{r}
df <- readxl::read_xlsx('../../../data/math-anxiety-raw-data.xlsx')
df_clean <- read.csv('../../../data/affectoutcome_norm_predict_1102.csv')
head(df)
head(df_clean)
```

### Data Clearning (raw data)

```{r}
tma_lst <- c("TMA_1", "TMA_2", "TMA_3","TMA_4", "TMA_5", "TMA_6",
             "TMA_sum", "TMA_avg")
df2 <- df %>% select(9,Condition, Sex, chicago, nonwhite, pretest,
              MW_day1_avg, MW_day2_avg, SI_avg, tma_lst,
              Understand_avg, Del_OverallAcc)
# Get rid of row that has NA values in Condition and/or Sex
df2 <- df2 %>% filter(Condition == 1 | Condition == 2)
df2 <- df2 %>% filter(!is.na(Sex))
df2$Condition[df2$Condition == 2] <- 0
# Assign 0 for "boy" and 1 for "girl"
df2$Sex[df2$Sex == 1] <- 0
df2$Sex[df2$Sex == 2] <- 1
# Make sure variable is categorical
df2$chicago <- as.numeric(df2$chicago)
df2$Sex <- as.numeric(df2$Sex)
df2$nonwhite <- as.numeric(df2$nonwhite)
df2$Condition <- as.numeric(df2$Condition)
```

Add MW average

```{r}
df2 <- df2 %>% mutate(MW_avg = rowMeans(df2[6:7], na.rm=TRUE))
head(df2[19])
```

### Data Cleaning (clean data)

The clean, imputed data provided contains `_sum` variable only for both `TMA`, `MW`, and `SI`. We want avaerage values for both math anxiety, mind wandering, and situational interest.

-   Trait math anxiety: 6 questions

-   Mind wandering: 12 questions

-   Situational interest: 14 questions

```{r}
df_clean <- df_clean %>% mutate(
  # average trait math anxiety
  TMA_avg = tma_sum_rv / 6,
  # average mind wandering
  MW_avg = ( mw_day1_sum + mw_day2_sum ) / 12,
  # average situational interest
  SI_avg = si_sum / 14
)
```

We also need to add students' accuracy scores to this dataset:

```{r}
colnames(df2)[1] <- "subjectid"
df3 <- df2 %>% select(1,19,2)
df_clean2 <- left_join(df_clean, df3, by = "subjectid") %>%   filter(!is.na(Del_OverallAcc))
```

```{r}
df_clean2n <- df_clean2
# standarize data
df_clean2[6:21] <- df_clean2[6:21] %>% mutate_all(~(scale(.) %>% as.vector))
head(df_clean2)
```

### Final Sample Size for Data Analysis/Modeling

#### Overall Descriptive

```{r}
skim(df_clean2)
```

Full sample size after preprocessing: $n = 224$

```{r}
control_n <- df_clean2 %>% 
  filter(Condition == 0) %>% group_by(chicago, gender) %>% count()
exp_n <- df_clean2 %>% 
  filter(Condition == 1) %>% group_by(chicago, gender) %>% count()
  
sample_size <- data.frame(rbind(t(control_n[3]),t(exp_n[3]))) %>% remove_rownames 
sample_size <- cbind(c("No Worked EX", "Worked EX"), sample_size)
sample_size <- sample_size %>% mutate(Total =  rowSums(sample_size[2:5])) %>% 
  adorn_totals("row") %>% data.frame()
colnames(sample_size) <- c("Condition", "Boy-I", "Girl-I", "Boy-C", "Girl-C", "Total")
sample_size
```

Total number of samples in data: $n = 224$

-   Number of samples in No Worked EX group: $n_{c} = 109$
    -   Girls: $n_{c,g} = 65$
    -   Boys: $n_{c,b} = 44$
-   Number of samples in Worked EX group: $n_{ex} = 115$
    -   Girls: $n_{c,g} = 64$
    -   Boys: $n_{c,b} = 51$

#### Ethnicity Descriptive

```{r}
# ethnicity descriptives
df_clean2 %>% group_by(race_rv) %>% count() %>% 
  mutate(p = n / nrow(df_clean2))
```

Summary:

-   21% white (48)
-   3% black (7)
-   18% Asian/Pacific Islander (41)
-   18% Hispanic/Latinx (41)
-   39% Mixed race/other (87)

## Gender Differences in TMA and MW

### Trait Math Anxiety

```{r}
cond <- c("No Worked EX", "Worked EX")
names(cond) <- c(0, 1)
location <- c("Irvine", "Chicago")
names(location) <- c(0, 1)
```

```{r}
# TMA mean by site, condition, and gender
mu <- df_clean2n %>% group_by(chicago, Condition,gender) %>% summarize(mean = mean(TMA_avg))
# math anxiety density plot
df_clean2n %>% ggplot(aes(x = TMA_avg, fill = as.factor(gender),
                   color = as.factor(gender))) +
  geom_density(alpha = 0.2) +
  theme_light() +
  geom_vline(data=mu, aes(xintercept=mean, color=as.factor
                          (gender)),linetype="dashed") +
  facet_grid(rows = vars(Condition),
             cols = vars(chicago),
             labeller = labeller(Condition = cond,
                                 chicago = location)) +
  scale_fill_manual(values = c("#FFC20A", "#0C7BDC"), labels = c("Boy", "Girl")) +
  scale_color_manual(values = c("#FFC20A", "#0C7BDC"), labels = c("Boy", "Girl")) +
  labs(x = "Math Anxiety", y = "Density",
       fill = "Gender", color = "Gender") +theme(legend.position="bottom")
```

Overall, all students seemed to have a low average math anxiety (1 \~ 2). In Irvine, girls seemed to have higher math average than boys in both condition. However, in Chicago, boys seemed to have higher average than girls.

Girls seemed to have a higher math anxiety compared to boys overall. Check if the difference is significant in each site by using ANOVA test:

```{r}
m1 <-  aov(TMA_avg ~ Condition + gender + chicago + 
             Condition*gender + Condition*gender*chicago,
           data = df_clean2)
summary(m1)
```

No significant difference between gender on math anxiety was found by gender.

### Mind Wandering

```{r}
# TMA mean by site, condition, and gender
mu <- df_clean2n %>% group_by(chicago, Condition,gender) %>% summarize(mean = mean(MW_avg))
# math anxiety density plot
df_clean2n %>% ggplot(aes(x = MW_avg, fill = as.factor(gender),
                   color = as.factor(gender))) +
  geom_density(alpha = 0.2) +
  theme_light() +
  geom_vline(data=mu, aes(xintercept=mean, color=as.factor
                          (gender)),linetype="dashed") +
  facet_grid(rows = vars(Condition),
             cols = vars(chicago),
             labeller = labeller(Condition = cond,
                                 chicago = location)) +
  scale_fill_manual(values = c("#FFC20A", "#0C7BDC"), labels = c("Boy", "Girl")) +
  scale_color_manual(values = c("#FFC20A", "#0C7BDC"), labels = c("Boy", "Girl")) +
  labs(x = "Mind Wandering", y = "Density",
       fill = "Gender", color = "Gender") +theme(legend.position="bottom")
```

In Irvine, both girls and boys seemed to have similar low average mind wandering (1\~2); worked example group showing somewhat higher mind wandering average than the worked example students in Irvine. On the other hand, in Chicago, boys had higher mind wandering than boys in both groups.

Check if the difference is significant in each site by using ANOVA test:

```{r}
m1 <-  aov(MW_avg ~ Condition + gender + chicago + 
             Condition*gender + Condition*gender*chicago,
           data = df_clean2)
summary(m1)
```

The results show that mind wandering was different between worked example and no worked example group, regardless of the cite location or gender. No gender difference was found in mind wandering.

```{r}
# plot the overall mind wandering average but by the condition
df_clean2n %>% group_by(Condition) %>% 
  summarize(mean = mean(MW_avg)) %>% 
  ggplot(aes(x = as.factor(Condition), y = mean)) +
  geom_bar(position = "dodge", stat = "identity", 
           color = "black", fill = "#ef476f", alpha = 0.5) +
  geom_text(aes(label=round(mean,2)),  
            position=position_dodge(width=0.9), 
            vjust=-0.25) +
  ylim(0,6) +
  labs(title = "Average Mind Wandering by Condition",
    x = "Condition", y = "Mind Wandeirng") + 
  scale_x_discrete(labels=c("No WE Group", "WE Group")) +
  theme_light()
```

WE group had a lower average mind wandering than no WE group.

## Modeling

### Overall Accuracy

**Model definition**

```{r}
m_acc <- '
# regressions
MW_avg ~ 1 + TMA_avg + SI_avg + gender + TMA_avg:gender
Del_OverallAcc ~ 1 + TMA_avg + MW_avg + SI_avg + gender + MW_avg:gender + TMA_avg:gender
'
```

**Implement model**

Fitting the model to entire dataset:

```{r}
fit_acc1 <- sem(m_acc, data = df_clean2, std.lv = T)
summary(fit_acc1)
```

Multigroup analysis with condition:

```{r}
fit_acc2 <- sem(m_acc, data = df_clean2, std.lv = T, group = "Condition")
summary(fit_acc2)
```

**WE group:**

-   Math anxiety had a significant positive effect on mind wandering

**No WE group:**

-   Math anxiety had a significant positive effect on mind wandering

-   Situational interest had a negative effect on mind wandering

-   Mind wandering had a significant negative effect on accuracy scores

Looking closely at the differences in groups

```{r}
fit_acc2.constrained <- sem(m_acc, df_clean2, group = "Condition", group.equal = c("intercepts", "regressions"))
anova(fit_acc2, fit_acc2.constrained)
```

Some paths in the model are different among groups -\> Looking closely at the differences in the effect of math anxiety on mind wandering between groups

```{r}
# redefine model
m_acc.constrained <- '
# regressions
MW_avg ~ 1 + c("b1", "b1") * TMA_avg + SI_avg + gender +  TMA_avg:gender
Del_OverallAcc ~ 1 + TMA_avg + MW_avg + SI_avg + gender + MW_avg:gender +  TMA_avg:gender
'
# fit model
fit_acc2.constrained2 <- 
  sem(m_acc.constrained, df_clean2, group = "Condition")
anova(fit_acc2, fit_acc2.constrained2)
```

No significant difference between worked ex groups vs no worked ex groups on the effect of math anxiety on mind wandering.

### Perceived Understanding

**Model definition**

```{r}
m_und <- '
# regressions
MW_avg ~ 1 + TMA_avg + SI_avg + gender + TMA_avg:gender
understand_avg ~ 1 + TMA_avg + MW_avg + SI_avg + gender + MW_avg:gender + TMA_avg:gender
'
```

**Implement model**

Fitting the model to entire dataset:

```{r}
fit_und1 <- sem(m_und, data = df_clean2, std.lv = T)
summary(fit_und1)
```

Multigroup analysis with condition:

```{r}
fit_und2 <- sem(m_und, data = df_clean2, std.lv = T, group = "Condition")
summary(fit_und2)
```

**WE group:**

-   Math anxiety had significant positive effect on mind wandering

-   Mind wandering had a significant negative effect on perceived understanding

-   Situational interest had a significant positive effect on perceived understanding

-   Interaction effect between mind wandering and gender was significant

    -   Boys: $\beta_{mw} = -0.537$

    -   Girls: $\beta_{mw} = -0.101$

    -\> Mind wandering impacted boys more than girls on their perceived understanding (girls pay attention more?)

**No WE group:**

-   Math anxiety had a significant positive effect on mind wandering

-   Situational interest had a significant negative effect on mind wandering

-   Mind wandering had a significant negative effect on perceived understanding

Finding model difference between groups:

```{r}
fit_und2.constrained <- sem(m_und, df_clean2, group = "Condition", 
    group.equal = c("intercepts", "regressions"))
anova(fit_und2, fit_und2.constrained)
```

Some paths in the model may vary among groups -\> Looking at the effect of mind wandering on perceived understanding

```{r}
# redefine model
m_und.constrained <- '
# regressions
MW_avg ~ 1 + TMA_avg + SI_avg + gender +  TMA_avg:gender
understand_avg ~ 1 + TMA_avg + c("b1", "b1") * MW_avg + SI_avg + gender + MW_avg:gender +  TMA_avg:gender
'
# fit model
fit_und2.constrained2 <- sem(m_und.constrained, df_clean2, group = "Condition")
anova(fit_und2, fit_und2.constrained2)
```

No significant difference was observed on the effect of mind wandering on perceived understanding.
