---
title: "Math Anxiety Initial Modeling"
author: "Emi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Tidyverse

```{r}
library(tidyverse) 
```

# Load clean data

```{r}
df <- read.csv("../../data/clean2.csv")
head(df)
glimpse(df)
```

# Modeling

1.    Normal regression model

2.    Path Analysis

For each model, we are using same dependent variable and list of independent variables:

**Dependent variable: **Mathematical anxiety `tma`

**Independent variables: **Gender (`gender`), worked example (`cond`), mind wandering (`mw1`, `mw2`), understanding (`under_avg`), high order thinking (`HOT`), situational interest (`si_sum`)

## Normal Regression Model

### Model 1.1

```{r}
m1_1 <- df %>% 
        lm(formula = tma ~ cond + mw1 + mw2 + under_avg + HOT + under_avg + si_sum + gender)
summary(m1_1)
```
Features that have significant effects on math anxiety (p < 0.05): `mw1` (positive effect: 0.158) and `si_sum` (negative effect: -0.058)

### Model 1.2: Take the average of mind wandering

```{r}
# Take the average of mind wandering day 1 and day 2
df <- df %>% 
        mutate( mw_avg = rowMeans(df[6:7], na.rm = TRUE))
```

```{r}
m1_2 <- df %>% 
        lm(formula = tma ~ cond + mw_avg + under_avg + HOT  + si_sum + gender)
summary(m1_2)
```

Features that have significant effects on math anxiety (p < 0.05): `mw_avg` (positive effect: 1.413) and `si_sum` (negative effect: -6.898e-02)


### Model 1.3: Take the ratio of day 1 and day 2

```{r}
# Take the ratio of mw1 and mw2
df <- df %>% 
        mutate(mw_ratio  = case_when(
          is.na(mw1) & is.na(mw2) ~ 0,
          is.na(mw1) | is.na(mw2) ~ 1,
          !is.na(mw1) & !is.na(mw2) ~ mw1 / mw2) )
```

```{r}
m1_3 <- df %>% 
        lm(formula = tma ~ cond + mw_ratio + under_avg + HOT + under_avg + si_sum + gender)
summary(m1_3)
```
Features that have significant effects on math anxiety (p < 0.05):  `si_sum` (negative effect: -0.08654)

### Model 1.4: Take the percent change of mw1 and mw2

$$mw_c = \frac{mw2 - mw1}{mw1}$$

```{r}
# Taking the percent change of mw1 and mw2
df <- df %>% 
        mutate(mw_ratio  = case_when(
          is.na(mw1) | is.na(mw2) ~ 0,
          !is.na(mw1) & !is.na(mw2) ~ (mw2 - mw1) / mw1) )
```

```{r}
m1_4 <- df %>% 
        lm(formula = tma ~ cond + mw_ratio + under_avg + HOT + under_avg + si_sum)
summary(m1_4)
```
Features that have significant effects on math anxiety (p < 0.05):  `si_sum` (negative effect: -0.08690)

### Model 1.5: Add interaction effects

We will consider worked example group as our interaction effects on the independent variables

Consider the following model:

`tma ~ cond + mw_avg + si_sum + mw_avg * cond + si_sum * cond + gender`

```{r}
m1_5a <- df %>% 
        lm(formula = tma ~ cond + mw_avg + si_sum + mw_avg * cond + si_sum * cond + gender)
summary(m1_5a)
```

`tma ~ cond + mw_avg + si_sum + mw_avg * gender + si_sum * gender + cond`

```{r}
m1_5b <- df %>% 
        lm(formula = tma ~ cond + mw_avg + si_sum + mw_avg * gender + si_sum * gender + cond)
summary(m1_5b)
```

`tma ~ cond + mw_avg + si_sum + mw_avg * cond + si_sum * cond + mw_avg * gender + si_sum * gender + mw_avg * gender * cond + si_sum * gender * cond`

```{r}
m1_5c <- df %>% 
        lm(formula = tma ~ cond + mw_avg + si_sum + mw_avg * cond + si_sum * cond + mw_avg * gender + si_sum * gender + mw_avg * gender * cond + si_sum * gender * cond)
summary(m1_5c)
```

`tma ~ cond + mw_avg + si_sum + mw_avg * cond + si_sum * cond + gender`

```{r}
m1_5bc <- df %>% 
        lm(formula = tma ~ cond + mw_avg + si_sum + mw_avg * cond + si_sum * cond + mw_avg * gender + si_sum * gender + mw_avg * gender * cond + si_sum * gender * cond)
summary(m1_5b)
```
Features that have significant effects on math anxiety (p < 0.05):  `mw_avg` (positive effect: 1.36529) `si_sum` (Negative effect: -0.11611), and `cond:si_sum` (small positive effect: 0.09319)

# Findings

*   Not enough evidence to suggest that worked examples have significant effects on students' mathematical anxiety levels

*   Taking the average of `mw1` and `mw2` resulted in better model compared to the ratio or percent change

*   Two features that had significant effects on math anxiety were: `mw_avg` and `si_sum`. Indicating that average mind wandering have positive impact on math anxiety and situational interest has negative impact on math anxiety

*   Very small effect on math anxiety by `cond:si_sum`, the interaction effect of worked example with situational interests. Situational interests may differ in each study group (small chance).

*   Switching to worked ex group from no-worked ex group had positive effect on students’ situational interests

# Visualizations

## Mind wanering average vs math anxiety

```{r}
df %>% 
  ggplot(aes(x = mw_avg, y = tma, color = as.factor(cond))) + 
  geom_point() + geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Mind Wandering (MW) Avg vs Math Anxiety (TMA)", x = "MW Avg", y = "TMA",
       color = "Condition") +
  scale_color_brewer(palette="Set1", labels = 
                      c("No Worked EX", "Worked EX")) + 
  theme_minimal()
```

## SI sum vs math anxiety

```{r}
df %>% 
  ggplot(aes(x = si_sum, y = tma, color = as.factor(cond))) + 
  geom_point() + geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Situational Interest (SI) vs Math Anxiety (TMA)", x = "SI", y = "TMA",
       color = "Condition") +
  scale_color_brewer(palette="Set1", labels = 
                      c("No Worked EX", "Worked EX")) + 
  theme_minimal()
```

# Modeling Independent Variables

```{r}
# Mind wandering
fit1 <- df %>% 
        lm(formula = mw_avg ~ cond + si_sum + under_avg + HOT)
summary(fit1)
```
```{r}
# Mind wandering
fit1 <- df %>% 
        lm(formula = si_sum ~ cond + mw_avg + under_avg + HOT)
summary(fit1)
```

## Findings

*   Mind wandering model: Situational interests (`si_sum`) and understanding level (`under_avg`) have significant effects (positive)

*   Situational interest model: Mind wandering (`mw_avg`) and understanding level (`under_avg`) have significant effects (positive)




