---
title: "Math Anxiety Initial Modeling"
author: "Emi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Packages

```{r}
library(lavaan)
library(semPlot)
library(OpenMx)
library(tidyverse)
library(knitr)
library(kableExtra)
library(GGally)
```

# Load clean data

```{r}
df <- read.csv("../../data/clean2.csv")
df <- df %>% 
        mutate( mw_avg = rowMeans(df[6:7], na.rm = TRUE))
head(df)
glimpse(df)
```

# Modeling

1.    Normal regression model: Simple regression model

2.    Path Analysis

For each model, we are using same dependent variable and list of independent variables:

**Dependent variable: **Mathematical anxiety `tma`

**Independent variables: **Gender (`gender`), worked example (`cond`), mind wandering (`mw1`, `mw2`), understanding (`under_avg`), high order thinking (`HOT`), situational interest (`si_sum`)

# Path Analysis

## Model 1

```{r}
# define model
model1 <- 'tma ~ cond + mw_avg + si_sum + under_avg + HOT + gender'
```
```{r}
# fit the model
fit1 <- cfa(model1, data = df)
summary(fit1)
```

```{r}
semPaths(fit1, 'std', layout = 'circle')
title("SEM Model 1",line=2.5)
```

## Model 2

```{r}
# define model
model2 <- '
tma ~ cond + mw_avg + si_sum + under_avg + HOT + gender
mw_avg ~ si_sum + under_avg
'
```
```{r}
# fit the model
fit2 <- cfa(model2, data = df)
summary(fit2)
```

```{r}
semPaths(fit2, 'std', layout = 'circle')
title("SEM Model 2",line=2.5)
```

## Model 3

```{r}
# define model
model3 <- '
tma ~ cond + mw_avg + si_sum + under_avg + HOT + gender
si_sum ~ mw_avg + under_avg
'
```
```{r}
# fit the model
fit3 <- cfa(model3, data = df)
summary(fit3)
```

```{r}
semPaths(fit3, 'std', layout = 'circle')
title("SEM Model 3",line=2.5)
```

# Findings

*   Still not enough evidence to suggest that conditions have significant effects on students' mathematical anxiety level

*   With SEM models, we can see the correlations between variables more clearly

*   Mind wandering and situational interests have significant effects on mathematical anxiety


