---
title: "Math COVID Anxiety Modeling 3"
author: "Emi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

# Load Dataset

```{r}
df <- read.csv("../../data/clean3.csv")
head(df)
```

# Model 1 

```{r}
fit1 <- df %>% 
        lm(formula = tma_avg ~ cond + mw_avg + understand_avg + si_avg + gender + hot)
summary(fit1)
```
Features that have significant effects on math anxiety (p < 0.05): `mw_avg` (positive effect: 0.060859) and `si_avg` (negative effect: -0.137749)

# Model 2: Interaction Effects (Condition)

```{r}
fit2 <- df %>% 
        lm(formula = tma_avg ~ cond + mw_avg + si_avg + mw_avg * cond + si_avg * cond + gender)
summary(fit2)
```
Features that have significant effects on math anxiety (p < 0.05): No features were proven to be statistically significant

# Model 3: Interaction Effects (gender)

```{r}
fit3 <- df %>% 
        lm(formula = tma_avg ~ cond + mw_avg + si_avg + mw_avg * gender + si_avg * gender + cond)
summary(fit3)
```
Features that have significant effects on math anxiety (p < 0.05): `mw_avg` (positive effect: 0.05264) and `si_avg` (negative effect: -0.20608)

# Model 4: Interaction Effects (gender and condition)

```{r}
fit4 <- df %>% 
        lm(formula = tma_avg ~ cond + mw_avg + si_avg + mw_avg * gender + si_avg * gender + mw_avg * cond + si_avg * cond + mw_avg * gender * cond 
           + si_avg * gender * cond)
summary(fit4)
```
Features that have significant effects on math anxiety (p < 0.05): `mw_avg` (positive effect: 0.124338) and `cond:mw_avg:gender` (positive effect: 0.098495)

Not really a good idea to include interaction effect in model

# Mind Wandering <-> Situational Interest

```{r}
df %>% ggplot(aes(x = si_avg, y = mw_avg, color = as.factor(cond))) + 
  geom_point() + geom_smooth(method = "lm", se = FALSE) + 
  theme_minimal() + 
  scale_color_brewer(palette="Set1", labels = 
                      c("No Worked EX", "Worked EX")) +
  labs(title = "Situational Interest (SI) vs Mind Wandering (MW) by Condition",
       x = "SI", y = "MW", color  = "Condition")

df %>% ggplot(aes(x = si_avg, y = mw_avg, color = as.factor(gender))) + 
  geom_point() + geom_smooth(method = "lm", se = FALSE) + 
  theme_minimal() + 
  scale_color_brewer(palette="Set1", labels = 
                      c("Girl", "Boy")) +
  labs(title = "Situational Interest (SI) vs Mind Wandering (MW) by Gender",
       x = "SI", y = "MW", color  = "Gender")

df %>% ggplot(aes(x = si_avg, y = mw_avg, color = tma_avg)) + 
  geom_point() + geom_smooth(method = "lm", se = FALSE) + 
  theme_minimal() + 
  labs(title = "Situational Interest (SI) vs Mind Wandering (MW) by TMA",
       x = "SI", y = "MW", color  = "TMA")
```

## Model 1: Situational Interest Model

```{r}
fit2_1 <- df %>% 
        lm(formula = si_avg ~ gender + cond + tma_avg + understand_avg + hot)
summary(fit2_1)
```
Features that have significant effects on situational interest (p < 0.05): `tma_avg` (negative effect: -0.23919) and `understand_avg` (positive effect: 0.01069)

## Model 2: Mind Wandering Model

```{r}
fit2_2 <- df %>% 
        lm(formula = mw_avg ~ gender + cond + tma_avg + understand_avg + si_avg + hot)
summary(fit2_2)
```
Features that have significant effects on situational interest (p < 0.05): `tma_avg` (positive effect: 1.65629) and `understand_avg` (negative effect: -0.07017)

# Findings

*   No evidence that showed condition had significant effect on neither `tma_avg`, `si_avg`, and `mw_avg`

*   `tma_avg` had significant negative effect on `si_avg` and `understanding_avg` had positive effect on `si_avg` -> lower math anxiety results in higher situational interest and higher understanding level results in higher situational interest

*   `tma_avg` had significant positive effect on `mw_avg` and `understanding_avg` had negative effect on `mw_avg` -> high math anxiety results in high mind wandering and high understanding results in low mind wandering

# Modeling with Emotions

```{r}
df %>% select(contains("enjoy"), contains("worry"))
```

## Features of Enjoy and Worry

Enjoy: `base_enjoy`, `enjoy_durimm_avg`,  `enjoy_del_avg`

Worry: `base_worry`, `worry_durimm_avg`, `worry_del_avg`

## Model 3: Percent change from base scores

New variables:

*  [**enjoy**] Change during the study: `enjoy_pc1` = (`enjoy_durimm_avg` - `base_enjoy`) / `base_enjoy`

*  [**enjoy**] Change after the study: `enjoy_pc2` = (`enjoy_del_avg` - `base_enjoy`) / `base_enjoy`

*  [**worry**] Change during the study: `worry_pc1` = (`worry_durimm_avg` - `base_worry`) / `base_worry`

*  [**worry**] Change after the study: `worry_pc2` = (`worry_del_avg` - `base_worry`) / `base_worry`

```{r}
# Create the variables
df <-  df %>% 
              mutate(enjoy_pc1 = (enjoy_durimm_avg - base_enjoy) / base_enjoy,
                     enjoy_pc2 = (enjoy_del_avg - base_enjoy) / base_enjoy,
                     worry_pc1 = (worry_durimm_avg - base_worry) / base_worry,
                     worry_pc2 = (worry_del_avg - base_worry) / base_worry)

df$enjoy_pc1[df$enjoy_pc1 == "NaN" | df$enjoy_pc1 == "Inf"] <- 0
df$enjoy_pc2[df$enjoy_pc2 == "NaN" | df$enjoy_pc2 == "Inf"] <- 0
df$worry_pc1[df$worry_pc1 == "NaN" | df$worry_pc1 == "Inf"] <- 0
df$worry_pc2[df$worry_pc2 == "NaN" | df$worry_pc2 == "Inf"] <- 0

head(df)
```

```{r}
# model with just percent change
# 1. All the percent change measures
fit3_1a <- df %>% 
        lm(formula = tma_avg ~ enjoy_pc1 + enjoy_pc2 + worry_pc1 + worry_pc2)
summary(fit3_1a)
# 2. Just using percent change using durimm scores
fit3_1b <- df %>% 
        lm(formula = tma_avg ~ enjoy_pc1 + worry_pc1)
summary(fit3_1b)
# 3. Using raw scores
fit3_1c <- df %>% 
        lm(formula = tma_avg ~ enjoy_durimm_avg + worry_durimm_avg + mw_avg + si_avg)
summary(fit3_1c)
```

```{r}
# model with just percent change
# mind wandering
fit3_1a <- df %>% 
  lm(formula = mw_avg~ enjoy_pc1 + worry_pc1 + tma_avg)
summary(fit3_2)
fit3_1b <- df %>% 
  lm(formula = mw_avg ~ enjoy_durimm_avg + worry_durimm_avg + tma_avg + si_avg)
summary(fit3_1b)
```

```{r}
# model with just percent change
# situational interest
fit3_2a <- df %>% 
  lm(formula = si_avg~ enjoy_pc1 + worry_pc1 + tma_avg)
summary(fit3_2a)
fit3_2b <- df %>% 
  lm(formula = si_avg~ enjoy_durimm_avg + worry_durimm_avg + tma_avg)
summary(fit3_2b)
```


