---
title: "Math Anxiety EDA 3"
format: html
editor: visual
---

# Load Libraries

```{r}
library(tidyverse)
```

# Load and Clean Dataset

```{r}
df <- readxl::read_xlsx('../../../../data/math-anxiety-raw-data.xlsx')
df <- df %>% filter(Condition == 1 | Condition == 2) %>% 
  drop_na(Sex)
df$Condition[df$Condition == 2] <- 0
df$Sex[df$Sex == 1] <- "boy"
df$Sex[df$Sex == 2] <- "girl"
df$chicago <- as.factor(df$chicago)
df$Sex <- as.factor(df$Sex)
```

# Reconsidering the Sampling of Data

What to look for:

1.  Number of students from Chicago/Irvine and Gender
2.  Students' prior knowledge about ratio before the study

## Number of Students from Chicago/Irvine

```{r}
df %>% group_by(chicago) %>% count()
```

Number of samples from Irvine: $n_i = 178$

Number of samples from Chicago: $n_c = 51$

```{r}
df %>% group_by(Sex) %>% count()
```

Number of girls in data: $n_g = 100$

Number of boys in data: $n_b = 129$

```{r}
df %>% group_by(Sex, chicago) %>% count()
```

Number of boys in Chicago: $n_{c,b} = 28$

Number of girls in Chicago: $n_{c,g} = 23$

Number of boys in Irvine: $n_{i,b} = 72$

Number of girls in Irvine: $n_{i,g} = 106$


## Students' prior knowledge about ratio before the study

```{r}
location <- c("Irvine", "Chicago")
names(location) <- c(0, 1)
gender <- c("Boy", "Girl")
names(gender) <- c("boy", "girl")
```

```{r}
df %>% group_by(chicago, Sex) %>% 
  summarize(avg_pretest = mean(pretest, na.rm = TRUE)) %>% 
  ggplot(aes(x =Sex, y = avg_pretest, fill = Sex)) + 
  geom_text(aes(label=round(avg_pretest, 3)),  position=position_dodge(width=0.9), 
            vjust=-0.25) +
  facet_grid(.~ chicago, 
             labeller = labeller(chicago = location)) +
  geom_bar(position = "dodge", stat = "identity", color = "black") +
  labs(title = "Avg Pretest Score by Gender and Location",
       x = "Gender", y = "Avg Pretest", fill = "Gener") +
  theme_minimal() + ylim(0,1)
```

Students in Irvine, both girls and boys, had higher pretest score average than the students in Chicago. In Chicago, boys had higher pretest score average than girls. 

```{r}
df %>% ggplot(aes(x = pretest, fill = Sex)) + 
  geom_bar(color = "black") +
  facet_grid(chicago ~ Sex,
             labeller = labeller(chicago = location,
                                 Sex = gender)) +
  theme_minimal() +
  labs(title = "Distribution of Pretest Score by Gender and Location",
       x = "Pretest", y = "Count")
```

Percentage of students who used ratio tactics during the pretest

```{r}
# Overall
prop.table(table(df$pretest))
```


```{r}
# Irvine
prop.table(table(df$pretest[df$chicago == 0]))
```

```{r}
# Chicago
prop.table(table(df$pretest[df$chicago == 1]))
```

68% of students in Irvine used ratio tactic during the pretest while 51% of students in Chicago used ratio tactic during the pretest. Overall, majority of students in Irvine already knew how to use the ratio tactic before the test.

**Question**: Did students who scored 1 in the `pretest` showed higher math anxiety than students who scored 0?

```{r}
df %>% filter(!is.na(pretest)) %>% 
  group_by(pretest, Sex) %>% summarize(mean(TMA_avg))
```

Not much of a difference in average math anxiety between students scored 0 and 1, and between different gender.

```{r}
# No worked example
df %>% filter(!is.na(pretest) & Condition == 0) %>% 
  group_by(pretest, Sex) %>% summarize(mean(TMA_avg))
```
In no worked example, there seems to be no difference in the average math anxiety between different pretest group and genders.

```{r}
# worked example
df %>% filter(!is.na(pretest) & Condition == 1) %>% 
  group_by(pretest, Sex) %>% summarize(mean(TMA_avg))
```
In worked example group, there seems to be no difference in the average math anxiety between different pretest group and genders, except for students who scored 0 in the pretest. Grils who scored 0 for the pretest had slightly higher average math anxiety than boys who scored 0.

### TMA question 5 average

```{r}
# No worked example
df %>% filter(!is.na(pretest) & Condition == 0) %>% 
  group_by(pretest, Sex) %>% summarize(mean(TMA_5))
```

```{r}
# Worked example
df %>% filter(!is.na(pretest) & Condition == 1) %>% 
  group_by(pretest, Sex) %>% summarize(mean(TMA_5))
```

Girls who scored 0 in pretest had higher average of math anxiety (quetion 5) than boys who scored 0 in pretest in the worked example group. The difference between girls and boys are higher for `TMA_5`