---
title: "Chromback's Alpha"
format: html
editor: visual
---

# Internal Consistency Reliability of Data

We are interested in looking at internal consistency reliability using Chromback's alpha for following variables:

-   Math anxiety scores (`TMA_1`, `TMA_2`, `TMA_3`, `TMA_4`, `TMA_5`, `TMA_6`)
-   Mind wandering scores (Day 1: `MW1_control_1`, `MW1_control_2`, `MW1_control_3`, `MW1_control_4`, `MW1_control_5`, `MW1_control_6`, `MW1_control_7`, Day 2: `MW2_control_1`, `MW2_control_2`, `MW2_control_3`, `MW2_control_4`, `MW2_control_5` )
-   Situational interest scores

## Load libraries

```{r}
library(tidyverse)
library(psych)
```

## Load data

```{r}
df <- readxl::read_xlsx('../../../data/math-anxiety-raw-data.xlsx') 
```

## Select variables of interest

```{r}
# Remove rows that had no condition assignment or gender
df1 <- df %>% filter(Condition == 1 | Condition == 2)
df1 <- df1 %>% filter(!is.na(Sex))
```

```{r}
# Select variables
tma <- df1 %>% select(TMA_1, TMA_2, TMA_3, TMA_4, TMA_5, TMA_6) %>% drop_na()
mw <- df1 %>% select(contains("mw_control")) %>% drop_na()
si <- df1 %>% select(contains("si_")) %>% select(1:14) %>% drop_na()
```

```{r}
# change col names of mw
col_names <- c("MW1_control_1", "MW1_control_2", "MW1_control_3", "MW1_control_4",
               "MW1_control_5", "MW1_control_6","MW1_control_7", "MW2_control_1",
               "MW2_control_2", "MW2_control_3", "MW2_control_4", "MW2_control_5")
colnames(mw) <- col_names
```

## Cromback's Alpha

Adequate internal consistency reliability $\geq .70$

Qualitative Descriptors of Cronbach's alpha

0.95 - 1.00 Excellent

.90 - .94 Great

.80 - .89 Good

.70 - .79 Acceptable

.60 - .69 Questionable

.00 - .59 Unacceptable

### TMA

```{r}
glimpse(tma)
```

Total observation of TMA: 229

```{r}
alpha(tma)
```

### Mind wandering

```{r}
glimpse(mw)
```

Total number of observations: 220

```{r}
alpha(mw)
```

### Situational Interests

```{r}
glimpse(si)
```

Total number of observations: 227

```{r}
alpha(si)
```

## Conclusion

All Cromback's alpha scores for TMA, MW, and SI measures were good. They all had good internal consistency reliability overall.
