---
title: "Math COVID EDA 3"
author: "Emi"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

# Load Dataset

```{r}
# Raw data
mathCOVID <- readxl::read_excel('../../data/KidsCOVID_Winter21_raw data.xlsx')
head(mathCOVID)
# Clean data
clean <- read.csv("../../data/affectoutcome_norm_predict_1102.csv")
```

# Clean Data Analysis

`base_enjoy`: How much are you enjoying this lesson right now?

`base_worry`: How anxious are you feeling right now?

`tma_sum_rv`: Mathematical anxiety

`mw_day1_sum` and `mw_day2_sum`: Mind wandering levels of day 1 and day 2

```{r}
glimpse(clean)
```

# Create a New Dataset

In a new dataset, contain: X, condition, race_rv, gender, chicago, tma_sum_rv, mw_day1_sum, mw_day2_sum, si_sum, understand_avg

```{r}
df <- mathCOVID %>% filter(is.na(drop)) %>% 
        select(9,Condition, race_rv, TMA_sum_rv, SI_sum, MW_day1_avg, MW_day2_avg, 
               Understand_avg, Base_Worry, Base_Enjoy, contains("durimm_avg"),
               contains("del_avg"), DurImm_HOT) 
df$Condition[df$Condition == 2] <- 0
col_names <- c("subjectid", "cond", "race", "tma", "si", "mw1", "mw2", "under_avg", 
               "worry0", "enjoy0", "worry1", "enjoy1", "worry2",
               "enjoy2", "HOT")
colnames(df) <- col_names
head(df)
````

```{r}
df2 <- inner_join(df, clean, by = "subjectid")
head(df2)
```

# Visualziations

```{r}
df %>% 
  ggplot(aes(x = tma, fill = cond)) + 
  geom_bar(position = "dodge") +
  theme_minimal() + 
  labs(title = "Mathematical Anxiety Distribution by Condition", 
       x = "Mathematical Anxiety (TMA)", y = "Count",
       fill = "Condition") +
  scale_fill_brewer(palette="Set1", labels = 
                      c("No Worked EX", "Worked EX"))

df %>% 
  ggplot(aes(x = tma, fill = cond)) + 
  geom_bar() + 
  facet_grid(.~cond) +
  scale_fill_brewer(palette="Set1", labels = 
                      c("No Worked EX", "Worked EX")) +
  labs(title = "Mathematical Anxiety Distribution by Condition",
       x = "TMA", y = "Count", fill  = "Condition")
```

```{r}
df %>% 
  ggplot(aes( x = tma, y = si, color = cond )) + 
  geom_point() + geom_smooth(method = "lm", se = FALSE) + 
  theme_minimal() + 
  scale_color_brewer(palette="Set1", labels = 
                      c("No Worked EX", "Worked EX")) +
  labs(title = "Mathematical Anxiety (TMA) vs Situational Interest (SI)",
       x = "TMA", y = "SI", color  = "Condition")
```

```{r}
df %>% 
  ggplot(aes( x = tma, y = mw1, color = cond )) + 
  geom_point() + geom_smooth(method = "lm", se = FALSE) + 
  theme_minimal() + 
  scale_color_brewer(palette="Set1", labels = 
                      c("No Worked EX", "Worked EX")) +
  labs(title = "Mathematical Anxiety (TMA) vs Mind Wandering (MW)",
       x = "TMA", y = "MW", color  = "Condition")
```

```{r}
df %>% 
  ggplot(aes( x = tma, y = under_avg, color = cond )) + 
  geom_point() + geom_smooth(method = "lm", se = FALSE) + 
  theme_minimal() + 
  scale_color_brewer(palette="Set1", labels = 
                      c("No Worked EX", "Worked EX")) +
  labs(title = "Mathematical Anxiety (TMA) vs Understanding Level",
       x = "TMA", y = "Understanding Level", color  = "Condition")
```


```{r}
df %>% 
  ggplot(aes(x = tma, y = HOT, color = cond)) + 
  geom_point() + geom_smooth(method = "lm", se = FALSE) + 
  theme_minimal() + 
  scale_color_brewer(palette="Set1", labels = 
                      c("No Worked EX", "Worked EX")) +
  labs(title = "Mathematical Anxiety (TMA) vs HOT",
       x = "TMA", y = "HOT", color  = "Condition")
```

# Implementing Possible Factors

## Possible Factors of Mathematical Anxiety

*   Student's background info: Gender, race, pre-worry and pre-enjoy scores

*   Condition: Worked example group vs no worked example group

*   Student's during/ability info: During and delayed worry/enjoy scores, mind wandeirng (day 1 and day 2), situational interest

```{r}
glimpse(df2)
write.csv(df2, "../../data/clean2.csv", row.names=FALSE)
```

```{r}
df2 %>% 
  ggplot(aes(x = tma, fill = as.factor(gender))) + 
  geom_histogram(color = "black",bins = 23)
```


