---
title: "Math Anxiety EDA 2"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(gridExtra)
library(ggpubr)
library(corrplot)
```

```{r}
df <- readxl::read_xlsx('../../../../data/math-anxiety-raw-data.xlsx')
```

## Look at the math anxiety for different gender

```{r}
df1 <- df %>% select(contains("tma"), chicago, Sex, Condition) %>% 
  na.omit() %>% filter(Condition == 1 | Condition == 2)
df1$Condition[df1$Condition == 2] <- 0
df1$Sex[df1$Sex == 1] <- "boy"
df1$Sex[df1$Sex == 2] <- "girl"
df1$chicago <- as.factor(df1$chicago)
df1$Sex <- as.factor(df1$Sex)
```

```{r}
df1 %>% group_by(Sex, Condition) %>% 
  summarize(tma_avg = mean(TMA_avg)) %>% 
  ggplot(aes(x = Condition, y = tma_avg, fill = Sex)) +
  geom_bar(position = "dodge", stat = "identity", color = "black") +
  geom_text(aes(label=round(tma_avg, 3)), position=position_dodge(width=0.9), 
            vjust=-0.25) +
  scale_x_discrete(labels=c("No Worked EX", "Worked EX")) +
  scale_fill_brewer(palette="Set1", labels = c("Boy", "Girl")) +
  theme_minimal() +
  labs("Average Comparison by Gender and Condition", x = "Condition",
       y = "TMA Avg") +
  ylim(0,5) 
```

```{r}
df1 %>% group_by(Sex) %>% 
  summarize(tma_avg = mean(TMA_avg)) %>% 
  ggplot(aes(x = Sex, y = tma_avg)) +
  geom_bar(position = "dodge", stat = "identity",
           fill = "darkgreen", color = "black") +
  geom_text(aes(label=round(tma_avg, 3)), position=position_dodge(width=0.9), 
            vjust=-0.25) +
  scale_x_discrete(labels=c("Boy", "Girl")) +
  theme_minimal() +
  labs("Average Comparison by Gender", x = "Gender",
       y = "TMA Avg") +
  ylim(0,5)
```

**Summary:** Girls have higher TMA average than boys overall and in both condition.

## Looking Closely at the TMA Question

Each question had score $\{ 1, 2, 3, 4, 5 \}$ where 1 being low stress and 5 being high stress. The survey measured students' **trait math anxiety**. Trait math anxiety refers to a trait of personality moment whereas state math anxiety refers to anxiety related to adverse situations in a specific moment.

1.  Having to complete a math worksheet by yourself
2.  Watching the teacher work out a math problem on the board
3.  Listening to the teacher talk for a long time in math
4.  Listening to another student in your class explain a math problem
5.  Starting a new topic in math
6.  Learning math online

```{r}
# Facet names for Condition and Gender
cond <- c("No Worked EX", "Worked EX")
names(cond) <- c(0, 1)
gender <- c("Boy", "Girl")
names(gender) <- c("boy", "girl")
```

```{r}
tma1 <- df1 %>% 
  ggplot(aes(x = TMA_1, fill = Sex)) +
  geom_bar(color = "black",position = "dodge") +
  facet_grid(.~Condition,
             labeller = labeller(Condition = cond)) +
  scale_fill_brewer(palette="Set1", labels = c("Boy", "Girl")) +
  labs(title = "TMA Question 1", x = "TMA Avg", y = "Count") +
  theme_minimal() 
tma2 <- df1 %>% 
  ggplot(aes(x = TMA_2, fill = Sex)) +
  geom_bar(color = "black",position = "dodge") +
  facet_grid(.~Condition,
             labeller = labeller(Condition = cond)) +
  scale_fill_brewer(palette="Set1", labels = c("Boy", "Girl")) +
  labs(title = "TMA Question 2", x = "TMA Avg", y = "Count") +
  theme_minimal() 
tma3 <- df1 %>% 
  ggplot(aes(x = TMA_3, fill = Sex)) +
  geom_bar(color = "black",position = "dodge") +
  facet_grid(.~Condition,
             labeller = labeller(Condition = cond)) +
  scale_fill_brewer(palette="Set1", labels = c("Boy", "Girl")) +
  labs(title = "TMA Question 3", x = "TMA Avg", y = "Count") +
  theme_minimal() 
tma4 <- df1 %>% 
  ggplot(aes(x = TMA_4, fill = Sex)) +
  geom_bar(color = "black",position = "dodge") +
  facet_grid(.~Condition,
             labeller = labeller(Condition = cond)) +
  scale_fill_brewer(palette="Set1", labels = c("Boy", "Girl")) +
  labs(title = "TMA Question 4", x = "TMA Avg", y = "Count") +
  theme_minimal() 
tma5 <- df1 %>% 
  ggplot(aes(x = TMA_5, fill = Sex)) +
  geom_bar(color = "black",position = "dodge") +
  facet_grid(.~Condition,
             labeller = labeller(Condition = cond)) +
  scale_fill_brewer(palette="Set1", labels = c("Boy", "Girl")) +
  labs(title = "TMA Question 5", x = "TMA Avg", y = "Count") +
  theme_minimal() 
tma6 <- df1 %>% 
  ggplot(aes(x = TMA_6, fill = Sex)) +
  facet_grid(.~Condition,
             labeller = labeller(Condition = cond)) +
  geom_bar(bins = 20, color = "black",
           position = "dodge") +
  scale_fill_brewer(palette="Set1", labels = 
                      c("Boy", "Girl")) +
  labs(title = "TMA Question 6", x = "TMA Avg", y = "Count") +
  theme_minimal() 
```

```{r}
tma_dist <- ggarrange(tma1, tma2, tma3, tma4, tma5, tma6, 
          common.legend = TRUE, legend="bottom", ncol = 3, nrow = 2)
tma_dist
```

```{r}
df1 %>% filter(TMA_2 > 3 & Condition == 1) %>% 
  count(Sex)
```

For TMA question 2, only boys in the worked example group showed anxiety scores 4 and 5.

## Overall Distribution of TMA by Gender and Condition

```{r}
df1 %>% 
  ggplot(aes(TMA_avg, fill = Sex)) +
  geom_histogram(bins = 13, color = "black") +
  facet_grid(Condition ~ Sex,
             labeller = labeller(Condition = cond,
                                 Sex = gender)) +
  theme_minimal() +
  labs(title = "TMA Avg Distribution by Gender and Condition", 
       x = "TMA Avg", y = "Count", fill = "Gender") +
   scale_fill_brewer(palette="Set1", labels = 
                      c("Boy", "Girl")) 
```

**Summary:** Girls have higher average in TMA compared to boys in both condition.

## Looking Closely at the Learning Achievements

```{r}
df2 <- df %>% select(Sex, Condition, pretest, chicago, contains("tma"),
                      DurImm_OverallAcc, Understand_avg) %>% 
        na.omit()
df2$Condition[df2$Condition == 2] <- 0
df2$Sex[df2$Sex == 1] <- "boy"
df2$Sex[df2$Sex == 2] <- "girl"
df2$Sex <- as.factor(df2$Sex)
df2$chicago <- as.factor(df2$chicago)
```

```{r}
df2 %>% ggplot(aes(x = DurImm_OverallAcc, fill = Sex)) +
  geom_histogram(bins = 13, color = "black") +
  facet_grid(Condition ~ Sex,
             labeller = labeller(Condition = cond,
                                 Sex = gender)) +
   theme_minimal() +
  labs(title = "Post Accuracy Avg by Gender and Condition", 
       x = "Accuracy Avg", y = "Count", fill = "Gender") +
   scale_fill_brewer(palette="Set1", labels = 
                      c("Boy", "Girl")) 
```

```{r}
df2 %>% ggplot(aes(x =  Understand_avg, fill = Sex)) +
  geom_histogram(bins = 17, color = "black") +
  facet_grid(Condition ~ Sex,
             labeller = labeller(Condition = cond,
                                 Sex = gender)) +
   theme_minimal() +
  labs(title = "Post Understanding Avg by Gender and Condition", 
       x = "Understanding Avg", y = "Count", fill = "Gender") +
   scale_fill_brewer(palette="Set1", labels = 
                      c("Boy", "Girl")) 
```

**Summary**: Not much of differences were observed in the distributions of understanding and accuracy scores of post-test between gender and condition.

## Correlation Plot

```{r}
glimpse(df2)
```

```{r}
res <- df2 %>% select(pretest, contains("TMA"), DurImm_OverallAcc, Understand_avg) %>% cor() %>% round(3)
```

```{r}
corrplot(res, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)
```

**Summary**: Different TMA measures are strongly correlated to each other. Students' understanding level was negatively correlated with some of the TMA measures including `TMA_1`,`TMA_6` , and `TMA_avg`.
