---
title: "Math Anxiety EDA"
format: html
editor: visual
---

## Worked Example and Math Anxiety

```{r}
library(tidyverse)
```

```{r}
df <- readxl::read_xlsx('../../../../data/math-anxiety-raw-data.xlsx')
```

# Look at the math anxiety for different gender

TMA Questionnaires:

1.  

```{r}
df %>%  filter(is.na(drop)) %>% 
  select(contains("sex"), contains("tma"), Condition) %>% 
  filter(!is.na(Sex) ) %>% 
  ggplot(aes(x = TMA_5, fill = Sex)) +
  facet_grid(.~Condition) +
  geom_bar(bins = 20, color = "black",
           position = "dodge") +
  scale_fill_brewer(palette="Set1", labels = 
                      c("Boy", "Girl")) +
  labs(x = "TMA Avg", y = "Count") +
  labs(title = "Distribution of TMA 5 by Gender and Condition") +
  theme_minimal() 
```

```{r}
df %>%  filter(is.na(drop)) %>% 
  select(contains("sex"), contains("tma"), Condition) %>% 
  filter(!is.na(Sex) ) %>% 
  ggplot(aes(x = TMA_6, fill = Sex)) +
  facet_grid(.~Condition) +
  geom_bar(bins = 20, color = "black",
           position = "dodge") +
  scale_fill_brewer(palette="Set1", labels = 
                      c("Boy", "Girl")) +
  labs(x = "TMA Avg", y = "Count") +
  theme_minimal() 
```

# Looking Closely at Participants' Background

Pre-test measured students' prior knowledge of ratio and proportional reasoning. If the students attempted any type of proportional strategy on the pretest item (one question), they received 1. Otherwise, they received 0. We will be comparing pre-test scores between different groups:

1.  Chicago schools vs. Irvine schools

2.  Boys vs. Girls

```{r}
df1 <- df %>% select(pretest, chicago, Sex, Condition) %>% 
  na.omit() 
df1$Condition[df1$Condition == 2] <- 0
df1$chicago <- as.factor(df1$chicago)
df1$Sex <- as.factor(df1$Sex)
```

## Looking at difference between Chicago and Irvine

```{r}
df1 %>% ggplot(aes(x = pretest, fill = chicago)) +
  geom_bar(position = "dodge") +
  scale_fill_brewer(palette="Set1", labels = 
                      c("Irvine", "Chicago")) +
  labs(title = "Pretest Scores by Schools", x = "Pretest Score", y = "Count", fill = "School") +
  theme_minimal()
```

```{r}
# Looking at the average pretest scores
pretest_school_plt <- df1 %>% group_by(chicago) %>% 
  summarize(pretest_avg = mean(pretest)) %>% 
  ggplot(aes(x = chicago, y = pretest_avg)) +
  geom_bar(position="dodge", stat="identity", 
           fill = "deepskyblue1", color = "black") +
  geom_text(aes(label=pretest_avg), position=position_dodge(width=0.9), 
            vjust=-0.25) +
  scale_x_discrete(labels=c("Irvine", "Chicago")) +
  labs(title = "Average Pretest Score by School", x = "School",
       y = "Pretest Avg") +
  ylim(0,1) +
  theme_minimal() + theme(plot.title = element_text(face = "bold", hjust = 0.5))
pretest_school_plt
```

```{r}
# Looking at the average pretest scores
pretest_sex_plt <- df1 %>% group_by(Sex) %>% 
  summarize(pretest_avg = mean(pretest)) %>% 
  ggplot(aes(x = Sex, y = pretest_avg)) +
  geom_bar(position="dodge", stat="identity", 
           fill = "cyan4", color = "black") +
  geom_text(aes(label=pretest_avg), position=position_dodge(width=0.9), 
            vjust=-0.25) +
  scale_x_discrete(labels=c("Girl", "Boy")) +
  labs(title = "Average Pretest Score by Gender", x = "Gender",
       y = "Pretest Avg") +
  ylim(0,1) +
  theme_minimal() + theme(plot.title = element_text(face = "bold", hjust = 0.5))
pretest_sex_plt
```

```{r}
pretest_plt <- ggarrange(pretest_school_plt, pretest_sex_plt, ncol = 2, nrow = 1)
ggsave(pretest_plt, file = "../../outputs/visualization/pretest.png")
```

**Summary**: Irvine schools had higher pretest scores than Chicago, which indicate that more students from Irvine knew how to approach ratio problem at the start of the experiment than the students from Chicago. On the other hand, there was not that much of a difference in pretest scores between girls an boys. Now, we look at the pretest score in the worked example group vs no worked example group.

```{r}
pretest_sex_cond <- df1 %>% group_by(Condition, Sex) %>% 
  summarize(pretest_avg = mean(pretest)) %>% 
  ggplot(aes(x = Condition, y = pretest_avg, fill = Sex)) +
  geom_bar(position="dodge", stat="identity", color = "black") +
  geom_text(aes(label=pretest_avg), position=position_dodge(width=0.9), 
            vjust=-0.25) +
  labs(title = "Average Pretest Score by Gender and Condition",
       x = "Condition", y = "Pretest Avg") +
  scale_fill_brewer(palette="Set1", labels = 
                      c("Girl", "Boy")) +
  ylim(0,1) + theme(plot.title = element_text(face = "bold", hjust = 0.5)) +
  scale_x_discrete(labels=c("No Worked EX", "Worked EX")) +
  theme_minimal()
pretest_sex_cond
```

```{r}
pretest_cond <- df1 %>% group_by(Condition) %>% 
  summarize(pretest_avg = mean(pretest)) %>% 
  ggplot(aes(x = Condition, y = pretest_avg)) +
  geom_bar(position="dodge", stat="identity", color = "black", fill = "cyan4")+
  geom_text(aes(label=pretest_avg), position=position_dodge(width=0.9), 
            vjust=-0.25) +
  ylim(0,1) + scale_x_discrete(labels=c("No Worked EX", "Worked EX")) +
  labs(title = "Average Pretest Score by Condition",
       x = "Condition", y = "Pretest Avg") +
  theme_minimal()
pretest_cond
```

```{r}
ggsave(pretest_cond, file = "../../../../outputs/visualization/pretest_cond.png")
ggsave(pretest_sex_cond, file = "../../../../outputs/visualization/pretest_sex_cond.png")
```

```{r}
df %>% group_by(chicago) %>% 
  count()
```
