---
title: "Modeling 3"
format: html
editor: visual
author: Emi Cervantes
---

# Modeling (Final PART I)

# Load Libraries

```{r}
library(tidyverse)
```

# Load Dataset

```{r}
df <- readxl::read_xlsx('../../../data/math-anxiety-raw-data.xlsx')
```

## Clean Dataset

```{r}
tma_lst <- c("TMA_1", "TMA_2", "TMA_3","TMA_4", "TMA_5", "TMA_6",
             "TMA_sum", "TMA_avg")
df1 <- df %>% select(Condition, Sex, chicago, nonwhite, pretest,
              MW_day1_avg, MW_day2_avg, SI_avg, tma_lst,
              Understand_avg, Del_OverallAcc)
# Get rid of row that has NA values in Condition and/or Sex
df1 <- df1 %>% filter(Condition == 1 | Condition == 2)
df1 <- df1 %>% filter(!is.na(Sex))
df1$Condition[df1$Condition == 2] <- 0
# Assign 0 for "boy" and 1 for "girl"
df1$Sex[df1$Sex == 1] <- 0
df1$Sex[df1$Sex == 2] <- 1
# Make sure variable is categorical
df1$chicago <- as.factor(df1$chicago)
df1$Sex <- as.factor(df1$Sex)
```

# Math Anxiety by Gender

```{r}
df1 %>% ggplot(aes(x = TMA_avg, fill = Sex,
                   color = Sex)) +
  geom_density(alpha = 0.5) +
  theme_minimal() +
  facet_grid(cols = vars(chicago),
             labeller = labeller(chicago = location)) +
  scale_fill_manual(values = c("#F0E442", "#56B4E9"), labels = c("Boy", "Girl")) +
  scale_color_manual(values = c("#F0E442", "#56B4E9"), labels = c("Boy", "Girl")) +
  labs(x = "Math Anxiety", y = "Density",
       fill = "Gender", color = "Gender")
```

Girls seemed to have a higher math anxiety compared to boys overall. Check if the difference is significant by using ANOVA test.

```{r}
m1 <- aov(TMA_avg ~ SI_avg + MW_day2_avg + chicago + Sex*chicago, 
          data = df1)
summary(m1)
```


```{r}
m2 <- aov(Del_OverallAcc ~ TMA_sum + chicago + Sex + Sex*Condition, data = df1, type = "III")
summary(m2)
```

No significant difference between gender on learning achievement and math anxiety was found by gender.

# Math Anxiety by Group and Gender

```{r}
cond <- c("No Worked EX", "Worked EX")
names(cond) <- c(0, 1)
location <- c("Irvine", "Chicago")
names(location) <- c(0, 1)
```


```{r}
df1 %>% ggplot(aes(x = TMA_avg, fill = Sex,
                   color = Sex)) +
  geom_density(alpha = 0.5) +
  theme_minimal() +
  facet_grid(rows = vars(Condition),
             cols = vars(chicago),
             labeller = labeller(Condition = cond,
                                 chicago = location)) +
  scale_fill_manual(values = c("#F0E442", "#56B4E9"), labels = c("Boy", "Girl")) +
  scale_color_manual(values = c("#F0E442", "#56B4E9"), labels = c("Boy", "Girl")) +
  labs(x = "Math Anxiety", y = "Density",
       fill = "Gender", color = "Gender")
```

```{r}
m3 <- aov(TMA_avg ~ SI_avg + MW_day2_avg + chicago + Sex*Condition
          + Sex*Condition*chicago, 
          data = df1)
summary(m3)
```

```{r}
m4 <- aov(Del_OverallAcc ~ TMA_avg + SI_avg + MW_day2_avg  +
            chicago + Sex + Sex*Condition + Sex*chicago*Condition, data = df1)
summary(m4)
```

## Linear Regression

```{r}
m5 <- TMA_avg ~ MW_day2_avg + SI_avg + chicago + Condition + chicago*Sex + chicago*Condition + Sex*Condition + chicago*Sex*Condition
fit5 <- lm(m5, data = df1)
summary(fit5)
```

```{r}
m6 <- Del_OverallAcc ~ TMA_avg + MW_day2_avg + SI_avg + chicago + Condition + Condition*Sex + Condition*TMA_avg + 
            Condition*TMA_avg*chicago + Condition*TMA_avg*chicago*Sex
fit6 <- lm(m6, data = df1)
summary(fit6)
```








